---
title: "ggrob_OriginalHomeworkCode_04"
author: "Gianna Grob"
date: "10/17/2019"
output: html_document
---

# Question 1

> Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines:

* Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().
* When conducting a two-sample test, it should be p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”, the same as in the use of x and y in the function t.test().
* The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.
* The function should contain a check for the rules of thumb we have talked about (n∗p>5 and n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings. If this is violated, the function should still complete but it should also print an appropriate warning message.
* The function should return a list containing the members Z (the test statistic), P (the appropriate p value), and CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.

```{r}
Z.prop.test <- function(p1, n1, p2 = NULL, n2 = NULL, p0, conf.level = 0.95, alternative = "two.sided")
{
#If statement for if p1 is 0, as well as testing for normality for one sample tests
  if(p1 == 0){
    return(0)}
    else {if((n1 * p1 < 5) | (n1 * (1-p1) < 5)){
          return(c("Warning: Distribution is not normal", (p1-p0) / sqrt((p0 * (1-p0))/n1)))}
          else(return((p1-p0) / sqrt((p0 * (1-p0))/n1)))}
}

#That worked! That gives us a statistic for a one sample test. Now I need to try to make it work for both one sample and two samples.

Z.prop.test.2 <- function(p1, n1, p2 = NULL, n2 = NULL, p0, conf.level = 0.95, alternative = "two.sided")
{
  if (p1 == 0){
    return(0)} #If p1 is 0, then 0 will be the statistic 
  
    else {if(is.null(p2) | is.null(n2)) #If p2 and n2 are null, test for the normal distribution of the one sample, and if normal, perform test
            if((n1 * p1 < 5) | (n1 * (1 - p1) < 5))
            return(c("Warning: Distribution is not normal", (p1-p0) / sqrt((p0 * (1-p0))/n1)))
            else(return((p1-p0) / sqrt((p0 * (1-p0))/n1)))} 
                else {if(!is.null(p2) | !is.null(n2))
                      if((n1 * p1 < 5) | (n1 * (1 - p1) < 5) | (n2 * p2 < 5) | (n2 * (1 - p2) < 5))
                      return(c("Warning: Distribution is not normal", (p1 - p2 - p0) / sqrt((p0 * (1-p0))/((1 / n1) + (1 / n2)))))   
                      else(return((p1 - p2 - p0) / sqrt((p0 * (1-p0))/((1 / n1) + (1 / n2)))))}     
}   

#I am having a lot of issues trying to get the second 'else' command to work. Is it possible to have two 'else' commands in a row without an 'if' statement in between them????

#Overall, I have not been very successful with this function. I am going to upload for original due time, but will work on this more and try to figure this out by Friday morning/afternoon.
```

Now I will test my function:

This is for when the distribution is not normal for one sample, a warning should be returned, but the function will still work.
```{r}
Z.prop.test(p1 = .33, n1 = 2, p0 = .60, p2 = NULL, n2 = NULL, alternative = "two.sided", conf.level = 0.95)
```

This is for when the distribtuion is normal, there should be no warning.
```{r}
Z.prop.test(p1 = .33, n1 = 100, p0 = .60, alternative = "two.sided", conf.level = 0.95)
```

# Question 2

> The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):

* Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).
* Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.
* Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.
* Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?
Looking at your two models, which do you think is better? Why?

```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall19/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE) #Load in our data
head(d) #See summary of data
```

Now that the data is loaded in, I will perform the regression analysis.

```{r}
library(ggplot2)

m <- lm(data = d, MaxLongevity_m ~ Brain_Size_Species_Mean)
summary(m) #Perform regression and see summary of results

g <- ggplot(data = d, aes(x = MaxLongevity_m, y = Brain_Size_Species_Mean))
g <- g + geom_point() 
g <- g + geom_smooth(method = "lm", se=FALSE, formula = y ~ x)
g <- g + geom_text(aes(200,400, label = ("y = β0 + β1Xi + ϵi"))) 
g #Plot data with regression line and written form of equation
```

Now, I will perform a T-test on the lm() data and and find the 90% confidence interval, as well as finding the prediction intervals. Then I will plot the new model.

```{r}
#Performing the T test
t <- coef(summary(m))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t

t$calct <- (t$Est - 0)/t$SE
t$calcp <- 2 * pt(t$calct, df = 212, lower.tail = FALSE)  # x2 because it is 2-tailed test
t

#Finding the 90% confidence level
ci <- confint(m, level = 0.90) 
ci

#Attempting different way of doing this to graph it better

df <- data.frame(cbind(d$MaxLongevity_m, d$Brain_Size_Species_Mean))
names(df) <- c("x", "y")
head(df)

ci <- predict(m, newdata = data.frame(MaxLongevity_m = d$MaxLongevity_m, Brain_Size_Species_Mean = d$Brain_Size_Species_Mean), interval = "confidence", 
    level = 0.90)  # for a vector of values
head(ci)

df <- cbind(df, ci)
names(df) <- c("x", "y", "CIfit", "CIlwr", "CIupr")
head(df)

#Finding prediction intervals

pi <- predict(m, newdata = data.frame(MaxLongevity_m = d$MaxLongevity_m, Brain_Size_Species_Mean = d$Brain_Size_Species_Mean), interval = "prediction", 
    level = 0.90)  # for a vector of values
head(pi)

df <- cbind(df, pi)
names(df) <- c("x", "y", "CIfit", "CIlwr", "CIupr", "PIfit", "PIlwr", 
    "PIupr")
df <- na.omit(df)
head(df)

#Plotting

e <- ggplot(data = df, aes(x = x, y = y))
e <- e + geom_point()
e <- e + geom_smooth(method = "lm", se=FALSE, formula = y ~ x)
e <- e + geom_line(aes(x = x, y = CIlwr), colour = "blue")
e <- e + geom_line(aes(x = x, y = CIupr), colour = "blue")
e <- e + geom_line(data = df, aes(x = x, y = PIlwr), colour = "red")
e <- e + geom_line(data = df, aes(x = x, y = PIupr), colour = "red")
e

#Wow! That totally did not work.... I need to find a different way to do this

#I don't know how to get these lines to fit with the data. Despite a lot of research and a deep look into the modules I am getting constantly stuck. I will wait for peer review for help.
```




